---
title: "Homework_6"
author: "Hannah Houts"
date: "March 4, 2019"
output: 
  html_document:
    keep_md: yes
---

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library(RColorBrewer)
options(scipen = 999)
#library(devtools)

```

## Resources
The idea for this assignment came from [Rebecca Barter's](http://www.rebeccabarter.com/blog/2017-11-17-ggplot2_tutorial/) ggplot tutorial so if you get lost go have a look. Please do not copy and paste her code!  

## Gapminder
For this assignment, we are going to use the dataset [gapminder](https://cran.r-project.org/web/packages/gapminder/index.html). Gapminder includes information about economics, population, and life expectancy from countries all over the world. You will need to install it before use.
```{r}
#install.packages("gapminder")
#install.packages("devtools")

devtools::install_github("cttobin/ggthemr")
```

```{r}
library("gapminder")
```

Please load the data into a new object called gapminder.
```{r}
gapminder <- 
  gapminder::gapminder
```

1. Explore the data using the various function you have learned. Is it tidy, are there any NA's, what are its dimensions, what are the column names, etc.



```{r}
summary(gapminder)
```

```{r}
head(gapminder)
```


_The data look tidy. Each column has one type of data, and all data are stored in columns. There dont appear to be any NAs in the data. There are 6 variables and there are 1704 observations_


2. We are interested in the relationship between per capita GDP and life expectancy; i.e. does having more money help you live longer on average. Make a quick plot below to visualize this relationship.
```{r}

gapminder %>%
ggplot(aes(x = gdpPercap, y = lifeExp))+
    geom_point()



```


3. There is extreme disparity in per capita GDP. Rescale the x axis to make this easier to interpret. How would you characterize the relationship?

```{r}

gapminder %>%
ggplot(aes(x = gdpPercap, y = lifeExp))+
    geom_point()+
    scale_y_log10()+
    scale_x_log10()



```

4. This should look pretty dense to you with significant overplotting. Try using a faceting approach to break this relationship down by year.

```{r}
gapminder %>%
ggplot(aes(x = gdpPercap, y = lifeExp))+
    geom_point()+
    scale_y_log10()+
    scale_x_log10()+
    facet_wrap(~year)

```

5. Simplify the comparison by comparing only 1952 and 2007. Can you come to any conclusions?

```{r}
gapminder %>%
    filter(year == 2007 | year == 1952)%>%
ggplot(aes(x = gdpPercap, y = lifeExp, color = year))+
    geom_point()+
    scale_y_log10()+
    scale_x_log10()

```

6. Let's stick with the 1952 and 2007 comparison but make some aesthetic adjustments. First try to color by continent and adjust the size of the points by population. Add `+ scale_size(range = c(0.1, 10), guide = "none")` as a layer to clean things up a bit.

```{r}

gapminder %>%
    filter(year == 2007 | year == 1952)%>%
ggplot(aes(x = gdpPercap, y = lifeExp, color = continent))+
    geom_point()+
    scale_y_log10()+
    scale_x_log10()+
    scale_size(range = c(0.1, 10), guide = "none")


```

7. Although we did not introduce them in lab, ggplot has a number of built-in themes that make things easier. I like the light theme for these data, but you can see lots of options. Apply one of these to your plot above.
```{r}
?theme_void



gapminder %>%
    filter(year == 2007 | year == 1952)%>%
ggplot(aes(x = gdpPercap, y = lifeExp, color = continent))+
    geom_point()+
    scale_y_log10()+
    scale_x_log10()+
    scale_size(range = c(0.1, 10), guide = "none")+
    theme_dark()



```



8. What is the population for all countries on the Asian continent in 2007? Show this as a barplot.

```{r}


gapminder %>%
    filter(continent == "Asia") %>%
    filter(year == 2007) %>%
    ggplot(aes(x = country, y = pop))+
    geom_col()+
    theme(axis.text.x = element_text(angle = 60, hjust=1))+                     scale_y_log10()
    
  

```

9. You should see that China's population is the largest with India a close second. Let's focus on China only. Make a plot that shows how population has changed over the years.

```{r}
gapminder %>%
    filter(country == "China") %>%
    ggplot(aes(x = year, y = pop))+
    geom_col()+
    theme(axis.text.x = element_text(angle = 60, hjust=1))
```

10. Let's compare China and India. Make a barplot that shows population growth by year using `position=dodge`. Apply a custom color theme using RColorBrewer.


```{r}
display.brewer.all()

```

```{r}
gapminder %>%
    filter(country == "China" | country == "India") %>%
    ggplot(aes(x = year, y = pop, fill = country))+
    geom_bar(position = "dodge", stat = "identity")+
    theme(axis.text.x = element_text(angle = 60, hjust=1))+
    theme_dark()+
    scale_fill_brewer(palette="PuRd")
```


## Push your final code to [GitHub](https://github.com/FRS417-DataScienceBiologists)
Make sure that you push your code into the appropriate folder. Also, be sure that you have check the `keep md` file in the knit preferences.